---
title: "<center>Seattle Pets: Explained<c/center>"
author: "Oleksiy Anokhin"
date: "March 26, 2019"
output: html_document
---

```{r, packages, echo = FALSE, message = FALSE, warning = FALSE}
# Install packages 
library(tidyverse)
library(ggthemes)
library(knitr)
library(leaflet)
library(leaflet.extras)
library(maps)
library(mapdata)
library(ggmap)
library(stringr)
library(viridis)
library(zipcode)
library(lubridate)
library(zoo)
# Disable scientific notation
options(scipen = 999)
```

_This small project is a part of a weekly [**#TidyTuesday**](https://github.com/rfordatascience/tidytuesday) challenge. This time I will explore ... Seattle pets._

_My first question was - How many pets per zip code in Seattle?_

_In order to figure it out, I would like to merge a given dataset with `data(zipcode)` from `zipcode` package, using `left_join()`. It will allow to assign each zipcode to a specific GPS location_ 

```{r, data, echo = FALSE, message = FALSE, warning = FALSE}
pets <- read_csv("seattle_pets.csv")
data(zipcode)
zips <- pets %>% group_by(zip_code) %>% summarize(number = n()) %>% arrange(desc(number))
seattle_zips <- left_join(zips, zipcode, by = c("zip_code"="zip"))
kable(head(seattle_zips))
```

_Then I will visualize number of pets per zip code._

<center>
```{r, map, echo = FALSE, message = FALSE, warning = FALSE}
my_popup <- paste0("<br><strong>Zip: </strong>", 
                   seattle_zips$zip_code,
                   "<br><strong>Number of pets per zip code: </strong>", 
                   seattle_zips$number)

leaflet(seattle_zips) %>% 
  setView(lng = -122.3, lat = 47.6, zoom = 11) %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addCircles(lng = ~longitude, lat = ~latitude, weight = 1,
             radius = ~number*0.2, color = "darkorange", opacity = 1, fillOpacity = 1, popup = ~my_popup)
```
</center>

_Then I would like to see if there any trends in owning cats vs dogs. I started from cats._

<center>
```{r, cat, echo = FALSE, message = FALSE, warning = FALSE}
cats <- pets %>% filter(species == "Cat")
cat_zips <- cats %>% group_by(zip_code) %>% summarize(number = n()) %>% arrange(desc(number))
cat_zips <- left_join(cat_zips, zipcode, by = c("zip_code"="zip"))
kable(head(cat_zips))

cat_popup <- paste0("<br><strong>Zip: </strong>", 
                   cat_zips$zip_code,
                   "<br><strong>Number of cats per zip code: </strong>", 
                   cat_zips$number)

leaflet(cat_zips) %>% 
  setView(lng = -122.3, lat = 47.6, zoom = 11) %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addCircles(lng = ~longitude, lat = ~latitude, weight = 1,
             radius = ~number*0.2, color = "yellow", opacity = 1, fillOpacity = 1, popup = ~cat_popup)
```
</center>

_Now let's explore dogs and map them._
<center>
```{r, dog, echo = FALSE, message = FALSE, warning = FALSE}
dogs <- pets %>% filter(species == "Dog")
dog_zips <- dogs %>% group_by(zip_code) %>% summarize(number = n()) %>% arrange(desc(number))
dog_zips <- left_join(dog_zips, zipcode, by = c("zip_code"="zip"))
kable(head(dog_zips))

dog_popup <- paste0("<br><strong>Zip: </strong>", 
                   dog_zips$zip_code,
                   "<br><strong>Number of dogs per zip code: </strong>", 
                   dog_zips$number)

leaflet(dog_zips) %>% 
  setView(lng = -122.3, lat = 47.6, zoom = 11) %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addCircles(lng = ~longitude, lat = ~latitude, weight = 1,
             radius = ~number*0.2, color = "yellow", opacity = 1, fillOpacity = 1, popup = ~dog_popup)
```
</center>

_I think it is obvious that people have more dogs in Seattle for some reason._

_Now I would like to see how the registration went by month._

_I start from converting data into more reliable format. From `chr` like `November 16 2018` to `int` like `Nov 2018`._

```{r, date, echo = FALSE, message = FALSE, warning = FALSE}
pets$license_issue_date <- mdy(pets$license_issue_date)
pets$license_issue_date <- as.yearmon(pets$license_issue_date, "%m/%Y")
pets_by_month <- pets %>% group_by(license_issue_date, species) %>% summarize(number = n())
kable(head(pets_by_month))
```

_The chart will look like this:_

```{r, date1, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(pets_by_month, aes(license_issue_date, number, color = species)) + geom_line(size = 1) +
theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Paired")
```

_As you can see, the active registering of pets in this dataset started after 2015, which raises questions what had happened before. I doubt that people did not have pets in Seattle before 2015? Maybe the city established new rules for such registration or something else?_

_Ok, let's check data from 2015 only. BTW, I realized that there are not only cats and dogs in this datasets, but also pigs and goats(!!!). I noticed it, when I visualized data on a previous chart._

```{r, date2, echo = FALSE, message = FALSE, warning = FALSE}
pets_by_month_2015 <- pets_by_month %>%
  filter(license_issue_date >= as.yearmon("Jan 2015") & license_issue_date <= as.yearmon("Dec 2018")) %>%
  filter(species == "Cat" | species == "Dog")
```

<center>
```{r, date3, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(pets_by_month_2015, aes(license_issue_date, number, color = species)) + geom_line(size = 1) +
theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
```
</center>

_As you can see, dogs prevail. More interesting that the pattern of registration is the same for cats and dogs. Not sure why though._ 

_Ok, everybody visualized cats and dogs and their popular names. I was curious about names of goats and pigs? Any pattern here?_

```{r, goats_pigs, echo = FALSE, message = FALSE, warning = FALSE}
goats_pigs <- pets %>%
  filter(species == "Pig" | species == "Goat") %>% group_by(animals_name, species) %>% summarize(number = n())
kable(goats_pigs)
```

_Which ones are your favorite? I love Brussels Sprout (goat), Coconut (pig), Othello (pig), Professor Nibblesworth (goat), and Sister Bertrille (pig). I was curious enough to google this Professor and found an [article](https://seattle.curbed.com/2017/12/15/16783234/seattle-popular-pet-names-2017), but not such person._ 

_Ok, let's at least visualize the most popular names of pets, as everybody did. I will filter by names with at least 100 times._

```{r, names, echo = FALSE, message = FALSE, warning = FALSE}
pets_names <- pets %>% group_by(species, animals_name) %>% summarize(number = n()) %>% arrange(desc(number)) %>% filter(number >= 100)
```

```{r, names1, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(pets_names, aes(reorder(animals_name, number), number, fill = species)) + geom_bar(stat = "identity", color  = "black") +
  coord_flip() +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
```

_It is interesting that so many cats are not registered with names. Why?_

**Conclusion**

_A pretty simple dataset, but still fun. I need a pig and I would call it Othello now._

**Author's Twitter**: [OleksiyAnokhin](https://twitter.com/OleksiyAnokhin)

