---
title: "<center>Visualizing 11 years of police traffic stops <br>in Albany, New York</center></br>"
author: "Oleksiy Anokhin"
date: "March 23, 2019"
output: html_document
---

```{r, packages, echo = FALSE, message = FALSE, warning = FALSE}
# Install packages 
library(tidyverse)
library(ggthemes)
library(knitr)
library(leaflet)
library(leaflet.extras)
# Disable scientific notation
options(scipen = 999)
```

_This small project is a part of a weekly [**#TidyTuesday**](https://github.com/rfordatascience/tidytuesday) challenge and uses [**The Stanford Open Policing Project**](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-19) data. More information about the project in this [video](https://www.youtube.com/watch?v=iwOWcuFjNfw)._

_I decided to analyze information about traffic stops by police in Albany, NY, because I lived in New York State before and was just curious about situation there. Albany data contains 10328 police stops of Albany drivers between December 2007 and December 2017._ 

_Firstly, I decided to see the number of stops by race. It immediately highlighted the first problem - there are too many NAs in this dataset._ 

```{r, race, echo = FALSE, message = FALSE, warning = FALSE}
albany_data <- read_rds("albany_data.rds")
albany_data <- rename(albany_data, age = subject_age, race = subject_race, sex = subject_sex)

# Filter by race
albany_stops_by_race <- albany_data %>% group_by(race) %>% summarise(number = n())
kable(albany_stops_by_race)
```

_In the form of a barchart stops by race will look like this:_ 

<center>
```{r, race1, echo = FALSE, message = FALSE, warning = FALSE}
# Visualize a number of stops by race
ggplot(albany_stops_by_race, aes(reorder(race, number), number, fill = race)) + geom_bar(stat = "identity", color  = "black") +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
```
</center>

_Then I asked myself about the age? Maybe drivers of a specific race are stopped differently based on their age? But boxplot looks similar. Moreover, it highlighted the second potential problem with data - outliers. Looks like there were drivers, who were 100+ yo, when they were stopped by police. Is it possible or it is a problem with typos?_ 

_In total I have 6 drivers with the age above 100, and 100 drivers with the age above 90._ 

<center>
```{r, age, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(albany_data, aes(race, age, fill = race)) + geom_boxplot() +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none")
```
</center>

_My third idea was to check how drivers are stopped by hour? Any specific hours with more stops?_ 
_Looks like that closer to midnight police stops drivers much more frequently and not very active early in the morning._ 
_Interesting that the majority of stops without the race identification were made during these late hours. Maybe policemen do not have time to fill in forms and skip some boxes in their forms?_

<center>
```{r, hour, echo = FALSE, message = FALSE, warning = FALSE}
albany_data$time <- strptime(albany_data$time, "%H:%M:%S")
albany_data$time <- as.integer(format(albany_data$time, "%H"))

ggplot(albany_data, aes(time, fill = race)) + geom_histogram(binwidth = 1, color = "black", alpha = 0.8) +
  scale_x_continuous(breaks = seq(0, 23, 1)) + 
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_blank()) +
  theme(legend.position = "bottom") + 
  ylim(0, 8000)
```
</center>

_Forth idea - any patterns with gender? The good thing - almost all drivers were idintified by their gender._ 

```{r, gender, echo = FALSE, message = FALSE, warning = FALSE}
albany_stops_by_gender <- albany_data %>% group_by(sex) %>% summarise(number = n())
kable(albany_stops_by_gender)
```

_Looks like stops of male drivers prevail significantly. Probably the majority of drivers are males. The chart looks like this:_

<center>
```{r, gender1, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(albany_stops_by_gender, aes(reorder(sex, number), number, fill = sex)) + geom_bar(stat = "identity", color  = "black") +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.position = "none") +
  ylim(0, 75000)
```
</center>

_Ok, maybe any specific patterns of race AND gender? Maybe some group was stopped disproportinally here? The table looks like this:_

```{r, gender2, echo = FALSE, message = FALSE, warning = FALSE}
albany_stops_by_race_gender <- albany_data %>% group_by(race, sex) %>% summarise(number = n())
kable(albany_stops_by_race_gender)
```

_We can see that NAs create problems here. Probably not too much sense to visualize this._ 

_My next question was about types of violations, which caused stops. Any patterns here?_

```{r, violation, echo = FALSE, message = FALSE, warning = FALSE}
albany_stops_by_violation_10 <- albany_data %>% group_by(violation) %>% summarise(number = n()) %>% arrange(desc(number)) %>% top_n(10)
kable(albany_stops_by_violation_10)
```

_Top-10 violations in Albany? Please see above. Violations of speed limits is the biggest problem. I am not a US citizen and I do not drive, so I am not sure how bad is this table in general._

_Ok, any patterns of violations by gender? Maybe groups have their "favorite" violations? I selected 20 the most frequent cases (not violations) with gender identification. And also created a stacked barchart for this table._ 

```{r, violation_gender, echo = FALSE, message = FALSE, warning = FALSE}
albany_stops_by_violation_sex <- albany_data %>% group_by(violation, sex) %>% summarise(number = n()) %>% arrange(desc(number)) 
albany_stops_by_violation_sex_20 <- head(albany_stops_by_violation_sex, 20)
kable(albany_stops_by_violation_sex_20)
```

<center>
```{r, violation_gender1, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(albany_stops_by_violation_sex_20, aes(reorder(violation, number), number, fill = sex)) + geom_bar(stat = "identity", color  = "black") + 
  coord_flip() +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_blank()) +
  ylim(0, 15000)
```
</center>

_Do you see any specific gender-based patterns? I do not really._ 

_Ok, my final attempt was to see any patterns in violations by race._

```{r, violation_race, echo = FALSE, message = FALSE, warning = FALSE}
albany_stops_by_violation_race <- albany_data %>% group_by(violation, race) %>% summarise(number = n()) %>% arrange(desc(number)) 
albany_stops_by_violation_race_20 <- head(albany_stops_by_violation_race, 20)
kable(albany_stops_by_violation_race_20)
```

_But with such data (full of NAs) it really does not make sense, as you can see from a table above, NAs prevail. As a result, I dropped all NAs from the `race` column and visualize top 20 cases._ 

```{r, violation_race1, echo = FALSE, message = FALSE, warning = FALSE}
albany_data_race_nas <- albany_data %>% filter_at(.vars = vars(race), .vars_predicate = any_vars(!is.na(.)))
albany_stops_by_violation_race_nas <- albany_data_race_nas %>% group_by(violation, race) %>% summarise(number = n()) %>% arrange(desc(number))
albany_stops_by_violation_race_nas_20 <- head(albany_stops_by_violation_race_nas, 20)
kable(albany_stops_by_violation_race_nas_20)

ggplot(albany_stops_by_violation_race_nas_20, aes(reorder(violation, number), number, fill = race)) + 
  geom_bar(stat = "identity", color  = "black") + 
  coord_flip() +
  theme_tufte() +
  theme(axis.title.y = element_blank()) +
  theme(axis.title.x = element_blank()) +
  theme(legend.title = element_blank())
```

_Finally, I decided to create the heatmap of violations, because just a map would have too much dots.
But again, the problem with data is NAs._

_Trying to build a heatmap, suing `leaflet.extras`, I got a warning message_

**In validateCoords(lng, lat, funcName) :**
  
  **Data contains 9851 rows with either missing or invalid lat/lon values and will be ignored**
  
_Decided to check total NAs:_

```{r, NAs, echo = FALSE, message = FALSE, warning = FALSE}
checked_nas <- sapply(albany_data, function(x) sum(is.na(x)))
kable(checked_nas)
```

_To avoid this I decided to remove NAs only from `lng` and `lat` colums (9851 rows) and visualize 93420 traffic police stops._ 

<center>
```{r, heatmap, echo = FALSE, message = FALSE, warning = FALSE}
albany_data_gps_nas <- albany_data %>% filter_at(.vars = vars(lng, lat), .vars_predicate = any_vars(!is.na(.)))

leaflet(albany_data_gps_nas) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addHeatmap(lng = ~lng, lat = ~lat,
             blur = 20, max = 0.05, radius = 15)
```
</center>

**Conclusion**

_Personally I did not notice any patterns, but I must admit that this data has problems with NAs and strange outliers in age._ 

**Author's Twitter**: [OleksiyAnokhin](https://twitter.com/OleksiyAnokhin)