---
title: "Learning About Wine From Scratch"
author: "Oleksiy Anokhin"
date: "6/17/2019"
output: html_document
---

```{r, packages, echo = FALSE, message = FALSE, warning = FALSE}
# Install packages 
library(tidyverse)
library(ggthemes)
```

**Introduction** 

I do not know anything about wine. A man supposed to know something, I guess. How to chose good wine and what
does "good" actually mean. I never cared about wine too much, even being born in a place, famous for its wines. Last Cristmas I just bought three bottles of Georgian wine, because I just expected it to be good + having some sympathy to local producers. 

This time I decided to learn about wine in a more scientific way. I picked a dataset from [Kaggle](https://www.kaggle.com/zynicide/wine-reviews), which stores information about almost 130000 different wines, its price, rankings, origin etc. 

```{r, data, echo = FALSE, message = FALSE, warning = FALSE}
#Read RDS data
wine_data <- readRDS("wine_data.rds")
```

This is how the dataset looks like. For me three the most interesting variables are: `country`, `points`, and `price`. These variables will allow me to make a quick analysis of wine quality and its price, taking into account its origin.  

```{r, data1, echo = FALSE, message = FALSE, warning = FALSE}
small_wine_data <- wine_data[1:1, ]
knitr::kable(head(small_wine_data))
```

Unfortunately, there many `NA` (missing data) for some variables, which will make the analysis a little less accurate. 

**Question 1. Visualize the correlation between price and ranking.**

Correlation always helpful to see the some basic patterns and outliers. 

```{r, data2, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(wine_data, aes(points, price)) + geom_point() + theme_tufte()
```
